---
title: "Calculating and visually communicating relationships among spatial data layers: isochrones"
author: "Kristy Henrich"
output: 
  html_document:
    theme: lumen
    toc: true
    toc_float: true
---
## Setup and load libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE, message = FALSE}
library(osmdata)
library(opentripplanner)
library(tidyverse)
library(sf)
library(ggthemes)
library(ggspatial)
library(sp)
library(stringr)
library(rgeos)
```

## Load locations
Load public schools in Boston

```{r, warning = FALSE, message = FALSE}
BOS_schools <- st_read(
  "http://bostonopendata-boston.opendata.arcgis.com/datasets/1d9509a8b2fd485d9ad471ba2fdb1f90_0.kml?outSR=%7B%22latestWkid%22%3A2249%2C%22wkid%22%3A102686%7D") 
```

## Get street data
Get street data for Boston, MA
```{r, warning = FALSE, message = FALSE}
opq(bbox = 'Boston MA USA') %>%
  add_osm_feature(key = 'highway') %>%
  osmdata_xml(file = 'OTP/graphs/default/boston_streets.osm')
```

```{r, warning = FALSE, message = FALSE}
MA_state_plane <- "+proj=lcc +lat_1=41.71666666666667 +lat_2=42.68333333333333 +lat_0=41 +lon_0=-71.5 +x_0=200000 +y_0=750000 +ellps=GRS80 +units=m +no_defs"

bos_street_features <- opq(bbox = 'Boston MA USA') %>%
  add_osm_feature(key = 'highway') %>%
  osmdata_sf()

bos_streets <- bos_street_features$osm_lines %>%
  st_transform(crs = MA_state_plane)
```

## Plot Boston streets

```{r, warning = FALSE, message = FALSE}
ggplot(bos_streets) +
  geom_sf() +
  theme_map()
```
## Set up Open Trip Planner

```{r, warning = FALSE, message = FALSE}
path_otp <- otp_dl_jar("OTP")
```

```{r, warning = FALSE, message = FALSE}
path_data <- file.path(getwd(), "OTP")
path_otp <- paste(path_data, "otp.jar",sep = "/")

otp_build_graph(otp = path_otp, dir = path_data, memory = 1024) 
```

```{r, warning = FALSE, message = FALSE}
otp_setup(otp = path_otp, dir = path_data, memory =1024)
```

# Connect to opentripplanner

```{r, warning = FALSE, message = FALSE}
otpcon <- otp_connect()
```

## Create isochrones

```{r, warning = FALSE, message = FALSE}
iso_5min_walk <- 
  otp_isochrone(otpcon = otpcon, fromPlace = BOS_schools, 
                mode = "WALK", cutoffSec = 300) %>%
  st_transform(crs = MA_state_plane) %>%
  mutate(mode = "walk")

iso_5min_drive <- 
  otp_isochrone(otpcon = otpcon, fromPlace = BOS_schools, 
                mode = "CAR", cutoffSec = 300) %>%
  st_transform(crs = MA_state_plane) %>%
  mutate(mode = "drive")

iso_all_modes <- rbind(iso_5min_drive, iso_5min_walk)

otp_stop()
```

## Figure 1
Map isochrones

I used Carole's tutorial to create this figure

```{r, warning = FALSE, message = FALSE}
right_side <- st_bbox(iso_all_modes)$xmax
left_side  <- st_bbox(iso_all_modes)$xmin
top_side <- st_bbox(iso_all_modes)$ymax
bottom_side <- st_bbox(iso_all_modes)$ymin

ggplot(iso_all_modes) +
  annotation_map_tile(zoomin = 0, type = "hotstyle", 
                      progress = "none") +
  geom_sf(aes(fill = mode), alpha = 0.5) +
  geom_sf(data = BOS_schools) +
  coord_sf(xlim = c(left_side, right_side), 
           ylim = c(bottom_side, top_side), expand = FALSE) +
  scale_fill_discrete(name = "Area that is reachable within 5 minutes",
                      labels = c("By car", "By foot"),
                      type = c("gray", "black")) +
  theme_map() +
  labs(caption = "Basemap Copyright OpenStreetMap contributors")
```

## Figure 2

Calculate the area of each isochrone and visualize the relationship between the size of a walkshed and the size of a driveshed via a scatter plot

Remove NULL data values

I used Carole's tutorial to create this figure

```{r, warning = FALSE, message = FALSE}
 iso_areas <- iso_all_modes %>%
   mutate(area = st_area(iso_all_modes)) %>%
   st_set_geometry(NULL) %>%
   pivot_wider(names_from = mode, values_from = area) 

iso_areas <- iso_areas %>% 
  filter(drive != "NULL", walk != "NULL") %>%
  filter(str_detect(drive,"c")==FALSE)

ggplot(iso_areas,
       aes(x = as.numeric(walk), y = as.numeric(drive))) +
  geom_point() +
  scale_x_continuous(name =
            "Area within a five-minute walking distance\nof a public school\n(square km)",
            breaks = breaks <- seq(10000, 300000, by = 20000),
            labels = breaks / 1000000) +
  scale_y_continuous(name =
            "Area within a five-minute driving distance\nof a public school\n(square km)",
            breaks = breaks <- seq(0, 2000000, by = 100000),
            labels = breaks / 1000000) +
  theme_bw()
```


## Figure 3

For figure 3, I was inspired by Whytne's maps and also used Alex's tutorial to look at public school accessibility across Boston by both walking and driving. I also switched to a 10 min walkshed and driveshed

Load Boston neighborhoods

```{r, warning = FALSE, message = FALSE}
nhoods <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/3525b0ee6e6b427f9aab5d0a1d0a1a28_0.kml?outSR=%7B%22latestWkid%22%3A2249%2C%22wkid%22%3A102686%7D", quiet = TRUE) 
```

Transform the Boston neighborhoods object into a spatial object
```{r, warning = FALSE, message = FALSE}
nhoods1 <- as(nhoods, "Spatial")

plot(nhoods1)
```

Create the many points around which will create isochrones

```{r, warning = FALSE, message = FALSE}
points <- spsample(nhoods1, n=1000, type='regular')

plot(points)
```

For the points above, make a dataframe so that can use in isochrone analysis below
```{r, warning = FALSE, message = FALSE}
points <- st_as_sf(x = points,                         
           coords = coords,
           crs = 4326)
```


### Figure 3A

Calculate the number of schools within a 10-minute walk for many points in Boston

```{r, warning = FALSE, message = FALSE}
otp_setup(otp = path_otp, dir = path_data, memory =1024)
```

```{r, warning = FALSE, message = FALSE}
otpcon <- otp_connect()
```

```{r}
iso_10min_walk <- 
  otp_isochrone(otpcon = otpcon, fromPlace = points, 
                mode = "WALK", cutoffSec = 600)
```


Count the number of schools in each 10min walkshed polygon
```{r, warning = FALSE, message = FALSE}
iso_10min_walk <- iso_10min_walk %>%
  mutate(walk_score = lengths(st_covers(geometry, BOS_schools)))
```

Extract coordinates and number of schools in the polygon so that can plot only points with schools
```{r, warning = FALSE, message = FALSE}
out <- data.frame(str_split_fixed(iso_10min_walk$fromPlace, ",", 2))

out <- st_as_sf(x = out,                         
           coords = c("X2", "X1"),
           crs = 4326)
  
out$walk_score <- iso_10min_walk$walk_score
```

Plot number of schools within a 10 min walk
```{r}
ggplot(nhoods) +
  geom_sf(fill="NA", color="grey")+
  geom_sf(data = out, aes(color=walk_score))+
  scale_color_gradientn(name="Number of public schools\nwithin a 10 min walk", colors=c("red", "yellow", "green"))+
  theme_map()+
  theme(legend.position = c(.7,0),
        plot.title = element_text(hjust = 0.5))+
  labs(title = "Public School Acessibility by Foot")
```




### Figure 3B

Calculate the number of schools within a ten-minute drive for many points in Boston

```{r, warning = FALSE, message = FALSE}
otp_setup(otp = path_otp, dir = path_data, memory =1024)
```

```{r, warning = FALSE, message = FALSE}
otpcon <- otp_connect()
```

```{r}
iso_10min_drive <- 
  otp_isochrone(otpcon = otpcon, fromPlace = points, 
                mode = "CAR", cutoffSec = 600)
```


Count the number of schools in each 10min driveshed polygon
```{r, warning = FALSE, message = FALSE}
iso_10min_drive <- iso_10min_drive %>%
  mutate(drive_score = lengths(st_covers(geometry, BOS_schools)))
```

Extract coordinates and number of schools in the polygon so that can plot only points with schools
```{r, warning = FALSE, message = FALSE}
out <- data.frame(str_split_fixed(iso_10min_drive$fromPlace, ",", 2))

out <- st_as_sf(x = out,                         
           coords = c("X2", "X1"),
           crs = 4326)
  
out$drive_score <- iso_10min_drive$drive_score
```

Plot number of schools within a 10 min drive
```{r}
ggplot(nhoods) +
  geom_sf(fill="NA", color="grey")+
  geom_sf(data = out, aes(color=drive_score))+
  scale_color_gradientn(name="Number of public schools\nwithin a 10 min drive", colors=c("red", "yellow", "green"))+
  theme_map()+
  theme(legend.position = c(.7,0),
        plot.title = element_text(hjust = 0.5))+
  labs(title = "Public School Acessibility by Car")
```



