---
title: "Layout two: using spatial distortion to make an argument"
author: "Kristy Henrich"
output: 
  html_document:
    theme: lumen
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries
For this assignment, I relied heavily on Carole's tutorial and Taelor's assignment. Alex inspired my use of some wonky projections and how to think about this in my coding, and Gemma also helped me reposition a legend on my map during our discussion on Monday, so thank you to Alex and Gemma!

```{r load libaries, results = FALSE, warning = FALSE}
library(ggplot2)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgeos)
library(tidyverse)
library(ggspatial)
library(cartogram)
library(ggthemes)
```

## South America equirectangular projection
```{r, results = FALSE, warning = FALSE}
sa_countries <- ne_countries(scale = 50, returnclass = "sf") %>%
  filter(continent == "South America")

sa_map <- ggplot(sa_countries) +
  geom_sf() +
  xlab("Longitude") + ylab("Latitude")
  ggtitle("Map of South America")
  theme_minimal()

sa_map + 
  annotation_scale()
```

## Define South America projections
Three projections:
  1. SR-ORG:7208: MONREF97: Mongolia
  2. Oblique Lambert azimuthal equal-area (appropriate projection according to spatialreference.org)
  3. EPSG:3003: Monte Mario / Italy zone 1
  
```{r, results = FALSE, warning = FALSE}
sa_mongolia = "+proj=tmerc +lat_0=0 +lon_0=105 +k=1 +x_0=500000 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
sa_ola = "+proj=laea +lon_0=-57.65625 +lat_0=-20.7777925 +datum=WGS84 +units=m +no_defs"
sa_italy = "+proj=tmerc +lat_0=0 +lon_0=9 +k=0.9996 +x_0=1500000 +y_0=0 +ellps=intl +units=m +no_defs"
```

### Projection of South America with Mongolia spatial reference
```{r, results = FALSE, warning = FALSE}
sa_map_mongolia <- ggplot(sa_countries) +
  geom_sf() + 
  coord_sf(crs = sa_mongolia) +
  theme_bw()

sa_map_mongolia
```

### Projection of South America with Oblique Lambert azimuthal equal-area 
This projection is more appropriate for South America than the other two selected 

```{r, results = FALSE, warning = FALSE}
sa_map_ola <- ggplot(sa_countries) +
  geom_sf() + 
  coord_sf(crs = sa_ola) +
  theme_bw()

sa_map_ola
```

### Projection of South America with Italy spatial reference
```{r, results = FALSE, warning = FALSE}
sa_map_italy <- ggplot(sa_countries) +
  geom_sf() + 
  coord_sf(crs = sa_italy) +
  theme_bw()

sa_map_italy
```

### Transform coordinate system to m (Oblique Lambert azimuthal equal-area)
```{r, results = FALSE, warning = FALSE}
sa_transformed <- sa_countries %>%
  st_transform(crs = sa_ola)
```

```{r, results = FALSE, warning = FALSE}
ggplot(sa_transformed) +
  geom_sf() +
  theme_bw()
```

## South America - percent of population in urban areas
For my variable data, I downloaded the World Bank's data on urban population as a percent of total population: https://data-worldbank-org.ezp-prod1.hul.harvard.edu/indicator/SP.URB.TOTL.IN.ZS

Throughout my assignment, I'll refer to my dataset as "% urban population (2019)"

```{r, results = FALSE, warning = FALSE}
sa_urban <- read_csv("C:/Users/krist/OneDrive - Harvard University/GitHub/kristy_henrich-vis/World_Bank_Percent_Urban_Population.csv") %>% 
               rename(country_name = "Country Name", urban_pop_19 = "2019")  %>% 
               select(country_name, urban_pop_19)

sa_countries_urban <- sa_transformed %>%
  left_join(sa_urban, by = c("name_sort" = "country_name")) %>%
  select(name_long, urban_pop_19)
```

### Plot % urban population as chloropleth map
```{r, results = FALSE, warning = FALSE}
sa_chloropleth <- ggplot(sa_countries_urban, aes(fill = urban_pop_19)) +
  geom_sf(color = NA) +
  scale_fill_viridis_c(
    name = "% urban population (2019)",
    breaks = seq(0, 100, by = 20),
    labels = formatC(seq(0, 100, by = 20), 
                     big.mark = ",", format = "f", digits = 0)) +
  theme_map() +
  theme(legend.background = element_blank())
```

### Save chloropleth map
```{r}
png(file="C:/Users/krist/OneDrive - Harvard University/GitHub/kristy_henrich-vis/SAchloropleth.png",
width=600, height=350)
sa_chloropleth
dev.off()
```

### Plot % urban population as non-continuous cartogram 
```{r, results = FALSE, warning = FALSE}
urban_cartogram_ncont <- sa_countries_urban %>%
  cartogram_ncont("urban_pop_19")

sa_cart_ncont <- ggplot(urban_cartogram_ncont, aes(fill = urban_pop_19)) +
  geom_sf(color = NA) +
  scale_fill_viridis_c(
    name = "% urban population (2019)",
    breaks = seq(0, 100, by = 20),
    labels = formatC(seq(0, 100, by = 20), 
                     big.mark = ",", format = "f", digits = 0)) +
  theme_map() +
  theme(legend.background = element_blank(),legend.position="right", 
        legend.title = element_text(size=9))
```
### Save non-contiguous cartogram map
```{r}
png(file="C:/Users/krist/OneDrive - Harvard University/GitHub/kristy_henrich-vis/SAncontcartogram.png",
width=600, height=350)
sa_cart_ncont
dev.off()
```


### Plot % urban population as proportional symbol map
```{r, results = FALSE, warning = FALSE}
urban_centeroids <- sa_countries_urban %>%
  st_centroid()

sa_centroids <- ggplot(sa_transformed) +
  geom_sf(fill = NA, color = "gray") +
  geom_sf(data = urban_centeroids, 
          aes(size = urban_pop_19),
          alpha = 0.5, color = "blue") +
  scale_size_continuous(name = "% urban population (2019)",
    breaks = seq(0, 100, by = 10),
    labels = formatC(seq(0, 100, by = 10), 
                     big.mark = ",", format = "f", digits = 0),
    range = c(0, 10)) +
  theme_void() 
```

### Save proportional symbol map
```{r}
png(file="C:/Users/krist/OneDrive - Harvard University/GitHub/kristy_henrich-vis/SAcentroids.png",
width=600, height=350)
sa_centroids
dev.off()
```

