---
title: "VIS Assignment 3"
author: "Kristy Henrich"
date: "9/22/2020"
output: html_document
---

## Setup & load libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE}
library(sf)
library(tidyverse)
library(ggthemes)
library(ggspatial)
library(units)
library(nngeo)
```

## Download spatial data
Data from data.boston.gov for:
  1. neighborhood boundaries (shape)  
  2. voting precincts (shape)  
  3. polling locations (point)  
  4. public libraries (point)
```{r}
nhoods <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/3525b0ee6e6b427f9aab5d0a1d0a1a28_0.geojson?outSR={%22latestWkid%22:2249,%22wkid%22:102686}", quiet = TRUE)

precincts <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/2bc185ec9b0b478d8c2f7e720d116863_0.geojson?outSR={%22latestWkid%22:2249,%22wkid%22:102686}", quiet = TRUE)

polls <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/f7c6dc9eb6b14463a3dd87451beba13f_5.kml?outSR={%22latestWkid%22:2249,%22wkid%22:102686}", quiet = TRUE)

lib <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/cb00f9248aa6404ab741071ca3806c0e_6.geojson?outSR={%22latestWkid%22:2249,%22wkid%22:102686}", quiet = TRUE)
```

## Transform spatial data
Transform data to the Massachusetts State Plane system using projection from spatialreference.org
```{r}
MA_state_plane <- "+proj=lcc +lat_1=41.71666666666667 +lat_2=42.68333333333333 +lat_0=41 +lon_0=-71.5 +x_0=200000 +y_0=750000 +ellps=GRS80 +units=m +no_defs"

nhoods <- nhoods %>%
  st_transform(MA_state_plane)

precincts <- precincts %>%
  st_transform(MA_state_plane)

polls <- polls %>%
  st_transform(MA_state_plane)

lib <- lib %>%
  st_transform(MA_state_plane)
```

## Quick map
```{r}
ggplot(nhoods) +
  geom_sf(fill = NA, color = "gray") +
  geom_sf(data = precincts, color = "lightblue") +
  geom_sf(data = polls, size = 0.01) +
  geom_sf(data = lib, color = "orange", size = 0.01) +
  theme_map() +
  annotation_scale()
```

## Figure 1. Number & proportion of A points within a specified distance of B points
Number and proportion of polling locations within 400 meters of a public library
```{r}
lib_buffer <- st_buffer(lib, dist = 400) %>%
  st_union()
ggplot(lib_buffer) +
  geom_sf() +
  theme_map()
```

```{r}
polls_lib <- polls[lib_buffer,]
  
ggplot(lib_buffer) +
  geom_sf() +
  geom_sf(data = polls_lib, 
          color = "forestgreen", 
          size = 1.5) +
  theme_map()
```

```{r}
polls <- polls %>%
  st_join(polls_lib) %>%
  mutate(by_lib = !is.na(Name.y))

n_polls_by_lib <- sum(polls$by_lib)
n_polls_by_lib
```

```{r}
n_polls <- length(polls$by_lib)

pct_polls_by_lib <- n_polls_by_lib / n_polls

pct_polls_by_lib
```


## Figure 2. Number and proportion of B points within C polygons
Calculate number and proportion of public libraries within neighborhoods

```{r}
nhoods <- nhoods %>%
  mutate(num_lib = lengths(st_covers(nhoods, lib)))
nhoods
```

```{r}
ggplot(nhoods) +
  geom_sf(color = NA, 
          aes(fill = num_lib)) +
  scale_fill_viridis_c(name = "Boston neighborhoods\nby number of public libraries") +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  
theme_map() +
  theme(legend.position = "bottom",
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```

## Figure 3. Number and proportion of C polygons containing A points
Number and proportion of neighborhoods containing polling locations

```{r}
nhoods <- nhoods %>%
  mutate(num_polls = lengths(st_covers(nhoods, polls)))

nhoods
```

```{r}
ggplot(nhoods) +
  geom_sf(color = NA, 
          aes(fill = num_polls)) +
  scale_fill_viridis_c(name = "Boston neighborhoods\nby number of polling locations") +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tl",
                         style = north_arrow_minimal()) +
  
theme_map() +
  theme(legend.position = "bottom",
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray")) 
```

## Figure 4. Number and proportion of D polygons that overlap with C polygons
Number and proportion of voting precincts that overlap with neighborhoods

```{r}
nhoods <- nhoods %>%
  mutate(num_precincts = lengths(st_overlaps(nhoods, precincts))) %>%
  mutate(has_precinct = num_precincts > 0)

nhoods
```

```{r}
n_prec_nhoods <- sum(nhoods$has_precinct)
n_prec_nhoods
```

```{r}
left_side  <- st_bbox(precincts)$xmin
top_side <- st_bbox(precincts)$ymax

ggplot(precincts) +
  geom_sf(fill = "lightblue", color = NA) +
  geom_sf(data = nhoods,
          aes(fill = has_precinct)) +
  scale_fill_manual(values = c("cornsilk1", "darkseagreen1"),
          name = "Boston neighborhoods\nby voting precincts", 
          labels = c("Neighborhood without\nvoting precincts",
                     "Neighborhood with \nvoting precincts")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  annotate(geom = "text", x = left_side, 
           y = top_side - 1000, 
           label = paste(n_prec_nhoods ,
                         "of Boston's", 
                         length(nhoods$Name),
                         "neighborhoods contain\nor overlap with", 
                         "voting precincts, \nmeaning they are fully covered by them."),
           hjust = 0, vjust = 0, size = 3) +
  theme_map() +
  theme(panel.background = element_rect(fill = "gray"),
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```

## Figure 5. Average number of A points in each D polygon
Average number of public libraries in each voting precinct

```{r}
precincts <- precincts %>%
  mutate(num_pub_lib = lengths(st_covers(precincts, lib)))

precincts
```
