---
title: "VIS Assignment 2"
author: "Kristy Henrich"
date: "9/18/2020"
output: 
  html_document:
    theme: lumen
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries
For this assignment, I relied heavily on Carole's tutorial and Taelor's assignment. I would like to be able to include country name labels on my maps to improve legibility, so I look forward to asking classmates for help on Monday. I also need to reposition one of my legends and would like help with this too

```{r load libaries, results = FALSE, warning = FALSE}
library(ggplot2)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgeos)
library(tidyverse)
library(ggspatial)
library(cartogram)
library(ggthemes)
```

## South America equirectangular projection
```{r, results = FALSE, warning = FALSE}
sa_countries <- ne_countries(scale = 50, returnclass = "sf") %>%
  filter(continent == "South America")

sa_map <- ggplot(sa_countries) +
  geom_sf() +
  xlab("Longitude") + ylab("Latitude")
  ggtitle("Map of South America")
  theme_minimal()

sa_map + 
  annotation_scale()
```

## Define South America projections
Three projections:
  1. SR-ORG:7208: MONREF97: Mongolia
  2. Oblique Lambert azimuthal equal-area (appropriate projection according to spatialreference.org)
  3. EPSG:3003: Monte Mario / Italy zone 1
  
```{r, results = FALSE, warning = FALSE}
sa_mongolia = "+proj=tmerc +lat_0=0 +lon_0=105 +k=1 +x_0=500000 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
sa_ola = "+proj=laea +lon_0=-57.65625 +lat_0=-20.7777925 +datum=WGS84 +units=m +no_defs"
sa_italy = "+proj=tmerc +lat_0=0 +lon_0=9 +k=0.9996 +x_0=1500000 +y_0=0 +ellps=intl +units=m +no_defs"
```

### Projection of South America with Mongolia spatial reference
```{r, results = FALSE, warning = FALSE}
sa_map_mongolia <- ggplot(sa_countries) +
  geom_sf() + 
  coord_sf(crs = sa_mongolia) +
  theme_bw()

sa_map_mongolia
```

### Projection of South America with Oblique Lambert azimuthal equal-area 
This projection is more appropriate for South America than the other two selected 

```{r, results = FALSE, warning = FALSE}
sa_map_ola <- ggplot(sa_countries) +
  geom_sf() + 
  coord_sf(crs = sa_ola) +
  theme_bw()

sa_map_ola
```

### Projection of South America with Greenland spatial reference
```{r, results = FALSE, warning = FALSE}
sa_map_italy <- ggplot(sa_countries) +
  geom_sf() + 
  coord_sf(crs = sa_italy) +
  theme_bw()

sa_map_italy
```

### Transform coordinate system to m (Oblique Lambert azimuthal equal-area)
```{r, results = FALSE, warning = FALSE}
sa_transformed <- sa_countries %>%
  st_transform(crs = sa_ola)
```

```{r, results = FALSE, warning = FALSE}
ggplot(sa_transformed) +
  geom_sf() +
  theme_bw()
```

## South America - percent of population in urban areas
For my variable data, I downloaded the World Bank's data on urban population as a percent of total population: https://data-worldbank-org.ezp-prod1.hul.harvard.edu/indicator/SP.URB.TOTL.IN.ZS

Throughout my assignment, I'll refer to my dataset as "% urban population (2019)"

```{r, results = FALSE, warning = FALSE}
sa_urban <- read_csv("C:/Users/Kristy Henrich/Documents/GitHub/kristy_henrich-vis/World_Bank_Percent_Urban_Population.csv") %>% 
               rename(country_name = "Country Name", urban_pop_19 = "2019")  %>% 
               select(country_name, urban_pop_19)

sa_countries_urban <- sa_transformed %>%
  left_join(sa_urban, by = c("name_sort" = "country_name")) %>%
  select(name_long, urban_pop_19)
```

### Plot % urban population as chloropleth map
```{r, results = FALSE, warning = FALSE}
ggplot(sa_countries_urban, aes(fill = urban_pop_19)) +
  geom_sf(color = NA) +
  scale_fill_viridis_c(
    name = "% urban population (2019)",
    breaks = seq(0, 100, by = 20),
    labels = formatC(seq(0, 100, by = 20), 
                     big.mark = ",", format = "f", digits = 0)) +
  theme_map() +
  theme(legend.background = element_blank())
```

### Plot % urban population as non-continuous cartogram 
```{r, results = FALSE, warning = FALSE}
urban_cartogram_ncont <- sa_countries_urban %>%
  cartogram_ncont("urban_pop_19")

ggplot(urban_cartogram_ncont, aes(fill = urban_pop_19)) +
  geom_sf(color = NA) +
  scale_fill_viridis_c(
    name = "% urban population (2019)",
    breaks = seq(0, 100, by = 20),
    labels = formatC(seq(0, 100, by = 20), 
                     big.mark = ",", format = "f", digits = 0)) +
  theme_map() +
  theme(legend.background = element_blank())
```

### Plot % urban population as proportional symbol map
```{r, results = FALSE, warning = FALSE}
urban_centeroids <- sa_countries_urban %>%
  st_centroid()

ggplot(sa_transformed) +
  geom_sf(fill = NA, color = "gray") +
  geom_sf(data = urban_centeroids, 
          aes(size = urban_pop_19),
          alpha = 0.5, color = "blue") +
  scale_size_continuous(name = "% urban population (2019)",
    breaks = seq(0, 100, by = 20),
    labels = formatC(seq(0, 100, by = 20), 
                     big.mark = ",", format = "f", digits = 0),
    range = c(0, 10)) +
  theme_void() 
```


